---
title: "Homework 5 — Choice 1: Mapping Denver Homicides"
author: "Tiffany Lipsey"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r libraries}
library(tidyverse)   # dplyr, ggplot2, readr, forcats
library(sf)          # simple features: spatial data as a data frame
library(tigris)      # downloads Census TIGER/Line boundary files

# Cache the downloaded boundaries so tigris doesn't re-download every knit
options(tigris_use_cache = TRUE)
```

## Step 1 — Read the data

```{r read}
homicides <- read_csv("./data/homicide-data.csv")

dim(homicides) 

```

## Step 2 - Filter for Denver and build two grouping variables
## Keep 3 most common race groups in Denver and put everything else into a single category

```{r denver}
denver <- homicides %>%
  filter(city == "Denver") %>%
  mutate(
    solved = if_else(disposition == "Closed by arrest", "Solved", "Unsolved"),
    solved = factor(solved, levels = c("Solved", "Unsolved")),
    victim_race = fct_lump_n(
      factor(victim_race),
      n = 3,
      other_level = "All other races"
    ),
    victim_race = fct_infreq(victim_race)          # order legend by frequency
  )

count(denver, solved)
count(denver, victim_race)

```

## Step 3 — Turn the points into an `sf` object

```{r points}
denver_sf <- st_as_sf(denver, coords = c("lon", "lat"), crs = 4326)

denver_sf
```

## Step 4 — Download tract boundaries with tigris

```{r tracts, results = "hide"}

options(tigris_use_cache = TRUE)

denver_tracts <- tracts(
  state  = "CO",
  county = "Denver",
  cb     = TRUE,  
  year   = 2010
)

```

## Step 5 Put both layers (sf from 3 and tracts from 4) in same coordinate system. Transform points into polygons CRS.

```{r crs}
st_crs(denver_tracts)$epsg   # 4269
st_crs(denver_sf)$epsg       # 4326

denver_sf <- st_transform(denver_sf, st_crs(denver_tracts))
```

## Step 6 — Build the map

```{r map, fig.width = 10, fig.height = 6}

denver_map <- ggplot() +
  geom_sf(
    data      = denver_tracts,
    fill      = "grey",
    color     = "black",
    linewidth = 0.3
  ) 

denver_map

```








