---
title: "Homework 5 — Choice 1: Mapping Denver Homicides"
author: "Tiffany Lipsey"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r libraries}
library(tidyverse)   # dplyr, ggplot2, readr, forcats
library(sf)          # simple features: spatial data as a data frame
library(tigris)      # downloads Census TIGER/Line boundary files

# Cache the downloaded boundaries so tigris doesn't re-download every knit
options(tigris_use_cache = TRUE)
```

## Step 1 — Read the data


```{r read}

homicides <- read_csv(
  here::here("data", "homicide-data.csv"),
  locale = locale(encoding = "Latin1")
)

dim(homicides) 

```

## Step 2 - Filter for Denver and build two grouping variables
## Keep 3 most common race groups in Denver and put everything else into a single category

```{r denver}
denver <- homicides %>%
  filter(city == "Denver") %>%
  mutate(
    solved = if_else(disposition == "Closed by arrest", "Solved", "Unsolved"),
    solved = factor(solved, levels = c("Solved", "Unsolved")),
    victim_race = fct_lump_n(
      factor(victim_race),
      n = 3,
      other_level = "All other races"
    ),
    victim_race = fct_infreq(victim_race)          # order legend by frequency
  )

count(denver, solved)
count(denver, victim_race)

```

## Step 3 — Turn the points into an `sf` object

```{r points}
denver_sf <- st_as_sf(denver, coords = c("lon", "lat"), crs = 4326)

denver_sf
```

## Step 4 — Download tract boundaries with tigris

```{r tracts, results = "hide"}

options(tigris_use_cache = TRUE)

denver_tracts <- tracts(
  state  = "CO",
  county = "Denver",
  cb     = TRUE,  
  year   = 2010
)

```

## Step 5 Put both layers (sf from 3 and tracts from 4) in same coordinate system. Transform points into polygons CRS.

```{r crs}
st_crs(denver_tracts)$epsg   # 4269
st_crs(denver_sf)$epsg       # 4326

denver_sf <- st_transform(denver_sf, st_crs(denver_tracts))
```

## Step 6 — Build the map

```{r map, fig.width = 10, fig.height = 6}
race_colors <- c(
  "Black"           = "#0072B2",   # blue
  "White"           = "#D55E00",   # vermillion
  "Hispanic"        = "#009E73",   # bluish green
  "All other races" = "#CC79A7"    # reddish purple
)

denver_map <- ggplot() +
  geom_sf(
    data      = denver_tracts,
    fill      = "lightblue",
    color     = "black",
    linewidth = 0.3
  ) +
  geom_sf(
    data  = denver_sf,
    aes(color = victim_race, shape = victim_race),
    size  = 1.8,
    alpha = 0.8
  )  +
  facet_wrap(~ solved) +
  scale_color_manual(values = race_colors, name = "Victim Race") +
  scale_shape_manual(values = c(16, 17, 15, 18), name = "Victim Race") +
  labs(
    title    = "Homicides in Denver, Colorado, 2010-2017",
    subtitle = "Solved (closed by arrest) versus unsolved, by victim race",
    caption  = paste(
      "Source: Washington Post homicide database.",
      "Tract boundaries: US Census Bureau via tigris."
    )
    ) +
  theme_void(base_size = 12) +
  theme(
  legend.position = "bottom",
  plot.title      = element_text(face = "bold", hjust = 0.5),
  plot.subtitle   = element_text(hjust = 0.5),
  plot.caption    = element_text(hjust = 0.5),
  strip.text      = element_text(face = "bold", size = 12,
                                 margin = margin(b = 6))
)

denver_map

```
##The large area to the right is the Denver airport. Good news! There were no homicides from 2010 to 2017 at the airport!

## Step 7 Save figure to figures/ directory

```{r save, eval = FALSE}

dir.create("figures")

ggsave(
  filename = here::here("figures", "denver-homicide-map.png"),
  plot     = denver_map,
  width    = 10,
  height   = 6,
  dpi      = 300
)
```








